\startSolutions{subspace}{Subspaces}
\begin{enumerate}[!HW!, start=1]
\itemspade It is not a subspace because $(0,0)\notin W$. Likewise, it is not closed under addition, e.g. $(1,0), (0,1)\in W$ but $(1,0)+(0,1)=(1,1)\notin W$. It is likewise not closed under scalar multiples, e.g. $(1,0) \in W$ but $2(1,0) = (2,0)\notin W$.
\itemspade It is not a subspace because it is not closed under addition, e.g. $(1,0), (0,1)\in W$ but $(1,0)+(0,1)=(1,1)\notin W$. Of course, there are some instances where the sum is contained, e.g. $(0.1, 0.2) + (-0.1, 0.3) = (0,0.5)\in W$, but closure under addition means that every possible sum is contained not just some sums.  It is likewise not closed under scalar multiples, e.g. $(1,0) \in W$ but $2(1,0) = (2,0)\notin W$. It does contain the zero vector.
\itemspade It is not a subspace because it is not closed under addition. Note that $(1,1), (2,4)\in W$ but $(1,1)+(2,4) = (3,5)\notin W$. It likewise fails closure under scalar multiples since $(1,1)\in W$ but $2(1,1) = (2,2)\notin W$. On the other hand, it does contain the zero vector.
\itemspade It is not a subspace because the sum of two points in $W$ need not belong to $W$, for example, $(1,0)+(0,1) = (1,1)\notin W$. On the other hand, $(0,0)\in W$ and $W$ is closed under scalar multiples. 
\itemspade It is not a subspace because it is not closed under scalar multiples, e.g. $(1,1)\in W$ but $(-1)(1,1) = (-1,-1)\notin W$. Is is closed under addition and contains the zero vector.
\itemspade It is not a subspace because the sum of two such functions will have a $y$-intercept of $2$, not $1$; so, the set is not closed under addition. Also, it is not closed under scalar multiples nor contains the zero function.
\item It is not a subspace because $(0,0)\notin W$. %Cyrus Kaveh
\itemspade It is a subspace, if $f(0)=0$ and $g(0)=0$ then $(f+g)(0) = f(0)+g(0) = 0+0=0$ and $(cf)(0) = c[f(0)]=c0=0$. Of course, the zero function is a member.
\itemspade It is not a subspace, because if the end behavior of $f$ is $\infty$ then the end behavior of $-f$ will be $\infty$; so, the set is not closed under scalar multiples. It likewise does not contain the zero function. It is closed under addition though.
\itemspade It is a subspace, because if $f(-x)=-f(x)$ and $g(-x)=-g(x)$ then $(f+g)(-x) = f(-x)+g(-x) = -f(x)-g(x) = -[f(x)+g(x)] = -(f+g)(x)$ and $(cf)(-x) = c[f(-x)] = c[-f(x)] = -[cf(x)] = -(cf)(x)$.  Of course, the zero function is a member.
\end{enumerate}

\vspace{-15 pt}